---
title: "Flu Occurence"
author: "Andriy Koval"
date: "January 27, 2020"
output: 
  html_document: 
    toc: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(magrittr)
library(dplyr)
library(ggplot2)
```

# Input data
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
path_file <- "../../data-public/raw/Tableau_10_Training_Files/Tableau 10 Training Practice Data.xlsx"
sheet_names <- readxl::excel_sheets(path_file)
dto <- list()
for(sheet_i in sheet_names){
   # i <- sheet_names[1]
  dto[[sheet_i]] <- readxl::read_xlsx(path_file, sheet = sheet_i)
}
ds0 <- dto$`05 - Flu Occurrence FY2013-2016` %>% tibble::as_tibble()
```

# Inspect data
Let us view the data in a dynamic table and see the key quantitative features of eahc column
```{r}
ds0 %>% DT::datatable()
ds0 %>% explore::describe_all()
```

# Tweak Data

Let us perform some basic wrangling to prepare data for graphing. We will  

1.  Rename the variables   

1.  Derive more date variables to assist with graph making   


```{r tweak-data}
names(ds0)
# let's store new names in a vector, in case we need to re-use it later (usually good idea)
new_varnames <- c(
  "date"
  ,"test_flu_positive_count"
  ,"total_specimen_count"
  ,"test_flu_positive_percent"
) 
ds1 <- ds0 
names(ds1) <- new_varnames

# let's create a few other date variables to enhance our analysis
ds1 <- ds1 %>% 
  dplyr::mutate(
    year        = lubridate::year(date)
    ,month      = lubridate::month(date)
    ,month_name = lubridate::month(date, label=TRUE) 
    # ,month_name = lubridate::month(date, label=TRUE, abbr=FALSE)
  )
ds1 %>% glimpse()
```

# Plan for today

Today we will proceed to make several graphs, improving the solution with each iteration.   

*   `g1` - will create the basic plot 

*   `g2` - will change the order of the months 

*   `g3` - will computed fiscal year from existing dates 

*   `g4` - will add custom colors and labels  

# `g1` Basic plot

```{r}
ds1 %>% glimpse(80)
g1 <- ds1 %>% 
  ggplot(aes(x = month, y = test_flu_positive_percent, color = factor(year)))+
  # ggplot(aes(x = month_name, y = test_flu_positive_percent, color = factor(year)))+
  geom_line()+
  # geom_line(aes(group = year))+
  theme_bw()
g1

```



# `g2` Order months

Let us change the order of months
```{r}
ds1 %>% glimpse(80)
month_order <- c(month.abb[10:12],month.abb[1:9])
ds2 <- ds1 %>% 
  dplyr::mutate(
    month_name = factor(month_name, levels = month_order )
  )

g2 <- ds2 %>% 
  ggplot(aes(x = month_name, y = test_flu_positive_percent, color = factor(year)))+
  geom_line(aes(group = year), size = 1.5)+
  theme_bw()
g2

# Built-in Constants:
# LETTERS
# letters
# month.abb
# month.name
# pi
```

# `g3` Fiscal year

```{r}
# let us create a column to store FISCAL year
# FISCAL year starts in October and ends in September
# if month is Oct, Nov, or Dec then they count toward the following year
# in other words:
# d <- ds1 %>% select(date,month_name,calendar_year = year)
# d %>% 
#   mutate(
#     move_to_next_year = (month_name %in% c("Oct","Nov","Dec"))
#     ,fiscal_year = calendar_year + move_to_next_year
#   )
# after we have done testing the logic, let's express it more succinctly
ds3 <- ds2 %>% 
  dplyr::mutate(
    fiscal_year = year + (month_name %in% c("Oct","Nov","Dec"))
  )
ds3
g3 <- ds3 %>% 
  ggplot(aes(x = month_name, y = test_flu_positive_percent, color = factor(fiscal_year)))+
  geom_line(aes(group = fiscal_year), size = 1.5)+
  theme_bw()
g3

```


# `g4` Colors and Lables 

```{r}
ds4 <- ds3 %>% 
  dplyr::mutate(
    y_display_1 = test_flu_positive_percent %>% scales::percent()
  )
ds4 %>% glimpse()
g4 <- ds4 %>% 
  ggplot(
    aes(
      x      = month_name
      ,y     = (test_flu_positive_percent *100) %>% round(digits =2)
      ,color = factor(fiscal_year)
      )
    )+
  geom_line(aes(group = fiscal_year), size = 1.5)+
  viridis::scale_color_viridis(discrete = TRUE, option = "magma", end = .8)+
  theme_bw()+
   labs(
    x  = "Month"
    ,y = "% Positive for Flu"
    , color = "Fiscal Year"
  )
g4
```


# Adding formats

```{r}
# let's demonstrate formatting into percentages
a <- c(.12345678, .0345 , .003456, .0001)
a
# method 1
a %>% round(4) %>%  scales::percent()
# method 2
(a*100) %>% round(4) %>% paste0("%")

```




# 

```{r eval = FALSE}
rmarkdown::render(
  input        = "../../analysis/lab_04_flu_occurrence/flu-occurrence.rmd"
  ,output_file = "../../analysis/lab_04_flu_occurrence/flu-occurrence.hmtl"
)
```













