---
title: "Preoperative Risk Factors"
author: "Andriy Koval"
date: "January 13, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Input the data

Let us reach into the Excel workbook and import the first data set. 

```{r}
ds <- readxl::read_excel(
  path = "Tableau 10 Training Practice Data.xlsx"
  , sheet = "01 - Preoperative Risk Factors"
)
print(ds)

```

Let us explore this data set

```{r}
dim(ds)
nrow(ds)
ncol(ds)
names(ds)
dplyr::glimpse(ds)

head(ds)
tail(ds)
```


Let us print the same table as in the Tableau for Healthcare textbook, chapter 4. 

```{r}
t4 <- dplyr::select(ds, "Preoperative Risk Factors","Hospital %","Comparison %")
t4 <- dplyr::filter( t4, !is.na(`Preoperative Risk Factors`) )
t4
```

Let us print this data set as a nicER table

```{r}
knitr::kable(t4,digits = 2, caption = "Table as seen in T4H-Chapter 4")
```





# Basics of R

> Everything is script and the environment is full of objects. 

The following is a brief summary of [R4DS-20-Vectors](https://r4ds.had.co.nz/vectors.html)

The key objects in R language are:  
- vectors  
- data frames  
- lists  
- functions  

## Vectors

A vector is a collection of values of the same type

```{r}
vector_numeric <- c(2, 5, 10)
vector_string  <- c("Joe", "Mary", "Bill")
vector_logical <- c(FALSE, TRUE, FALSE)

```

## Data Frames

A data frame is a rectangular object in which columns are vectors

```{r}
df <- data.frame(
  age     = c(2, 5, 10)
  ,name   = c("Joe", "Mary", "Bill")
  ,female = c(FALSE, TRUE, FALSE)
)
df
# alternatively 
ds <- tibble::tibble(
    age   = c(2, 5, 10)
  ,name   = c("Joe", "Mary", "Bill")
  ,female = c(FALSE, TRUE, FALSE)
)
ds

```
Compare the classes
```{r}
class(df)
class(ds)
```

A tibble is "an opinionated data frame", which behaves like a list. (see more at [R4DS-10-Tibbles](https://r4ds.had.co.nz/tibbles.html))

## Lists

A list is a collection of  vectors, data frames, or lists

```{r}
ls_vectors <- list(vector_numeric, vector_string, vector_logical)
(a <- ls_vectors[1]) # as a list
(b <- ls_vectors[[1]]) # as a vector
names(ls_vectors) <- c("age","name","female") # re-run above two lines to compare
# a better way of naming elements in the list
ls_vectors <- list(
  "age"     = vector_numeric
  ,"name"   = vector_string
  ,"female" = vector_logical
  )
ls_vectors
# still can refer

```

## Functions

Functions are instructions on what to do with other objects. There are "verbs" in the grammar of R programming. See more in [R4DS-19-Functions](https://r4ds.had.co.nz/functions.html). 

```{r}
# define function
compute_mean_of_a_vector <- function(x){mean(x)}
# how to use
compute_mean_of_a_vector(x = vector_numeric)
compute_mean_of_a_vector(x = vector_string)
compute_mean_of_a_vector(x = vector_logical)


# define function
compute_mean_of_a_column_in_a_dataset <- function(d,column_name){
  # d <- ds
  # column_name <- "age"
  # isolate the vector
  # vector_x <- as.vector(ds$age)
  # vector_x <- as.numeric(d[["age"]]) # compare with ds["age"]
  vector_x <- as.numeric(d[[column_name]]) # compare with ds["age"]
  # class(vector)
  (output <- mean(vector_x))
  return(output)
}
# how to use
compute_mean_of_a_column_in_a_dataset(d = ds, column_name = "age")
```
















