---
title: "Functions"
author: "Andriy Koval"
date: "2/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm( list = ls(all = TRUE)) # to clean the environment
cat("\014") # to clean the console

```


```{r message=FALSE, warning=FALSE}
library(magrittr) # pipes
library(dplyr)    # wrangling
library(ggplot2)  # graphing
library(tidyr)    # pivoting

```

# Simulate data

a) Simulate data for performance on three exams with the following specifications:  
- There are the total of 100 students in the class
- Exam 1 has the mean of 60 and standard deviation of 7
- Exam 2 has the mean of 73 and standard deviation of 9
- Exam 3 has the mean of 80 and standard deviation of 12
- There is a 7 % chance for a student to miss any of the exams

b) create a histogram featuring all three exams faceted vertically (hist: same x axis)  


```{r message=FALSE, warning=FALSE}
set.seed(42)
class_size <- 100
ds <- tibble::tibble(
  student_id = 1:class_size
  ,"exam1" = rnorm( n = class_size, mean = 60, sd = 7 )
  ,"exam2" = rnorm( n = class_size, mean = 73, sd = 9 )
  ,"exam3" = rnorm( n = class_size, mean = 80, sd = 12 )
  ,"exam1_missed" = runif(n = class_size ) < .07
  ,"exam2_missed" = runif(n = class_size ) < .07
  ,"exam3_missed" = runif(n = class_size ) < .07
) %>%
  dplyr::mutate(
    exam1  = ifelse(exam1_missed, NA, exam1)
    ,exam2 = ifelse(exam2_missed, NA, exam2)
    ,exam3 = ifelse(exam3_missed, NA, exam3)
  ) %>%
  dplyr::select(-exam1_missed, -exam2_missed, -exam3_missed)

ds %>%
  tidyr::pivot_longer(
    cols = c("exam1", "exam2", "exam3")
    ,names_to = "exam"
    ,values_to = "score"
  ) %>%
  ggplot( aes(x = score, fill = exam)) +
  geom_histogram(alpha = .3)+
  facet_grid(exam ~ .)+
  theme_bw()
```
# Compute total score

c) compute the total score for each student.

Hints:    
- you will need to use perform rowwise computation using either `rowwise` or `base::rowSums`  
- don't forget about `NA` values, each student must have a total score  

```{r}
ds %>% glimpse()

# solution 1
ds1 <- ds %>% 
  rowwise %>% 
  dplyr::mutate(
    total_score = sum(exam1, exam2, exam3, na.rm = TRUE)
  )
ds1
# solution 2
ds2 <- ds %>% 
  dplyr::mutate(
    total_score = rowSums(ds %>% select(exam1, exam2, exam3), na.rm = TRUE)
  )

ds2
```

# Design the graph

We would like to design the graph that would plot the progress of any given student.   
Hints:    
- start by subsetting data for a single student  
- You will need to isolate the values of the total score and compute the number of missing exams, store each as a numeric vector of length 1  
- when you have 1x1 tibble, a good trick to convert it to a vector of legnth 1 is to add `%>% as.list() %>% unlist() %>% as.numeric()` at the end of the expression  
- you will need to `pivot_longer` in order to make your graph  
- you will need to use `geom_point`, `geom_line`, and `geom_text`  
- don't forget about `aes(group = )` in your `geom_line`  

```{r message=FALSE, warning=FALSE}
selected_student_id = 1

d <- ds1 %>%  
  filter(student_id == selected_student_id ) %>% 
  tidyr::pivot_longer(c(exam1, exam2, exam3, total_score), names_to = "assessment", values_to = "score") 
d

selected_total_score <- d %>% filter(assessment == "total_score") %>% select(score) %>% as.list() %>% unlist() %>% as.numeric()
n_exams_missing <- d %>% filter(!assessment == "total_score") %>% summarize( n_missing = sum(is.na(score))) %>% as.list() %>% unlist() %>% as.numeric()
  
d %>% 
  filter(!assessment == "total_score") %>% 
  ggplot(aes(x = assessment, y = score  ))+
  geom_point(shape = 21, fill = "salmon", color = "blue", size = 4)+
  geom_line(aes(group = student_id), linetype = "dashed")+
  geom_text(aes(label = score %>% round(1)), hjust = -1)+
  labs(
    title = paste0(
      "Course performance for student [", selected_student_id,"]","\n",
      "Number of exams missing = ", n_exams_missing, "; Total score = ", selected_total_score %>% round(1)
    )
  )+
  theme_bw()
```

# Compose a function to plot one student

Create a function called `graph_student_progress` to plot all three exam scores. 
Hints:
- Your function should have two arguments: `d` and `selected_student_id`

```{r message=FALSE, warning=FALSE}
graph_student_progress <- function(d, selected_student_id){
  # subset data to focus on the relevant student
  d1 <- d %>%  
    filter(student_id == selected_student_id ) %>% 
    tidyr::pivot_longer(c(exam1, exam2, exam3, total_score), names_to = "assessment", values_to = "score") 
  # extract a value of the total score to use in the title
  selected_total_score <- d1 %>% filter(assessment == "total_score") %>% select(score) %>% 
    as.list() %>% unlist() %>% as.numeric()
  # extact a value of the number of missing exams to use in the title
  n_exams_missing <- d1 %>% filter(!assessment == "total_score") %>% summarize( n_missing = sum(is.na(score))) %>% 
    as.list() %>% unlist() %>% as.numeric()
  # plot the graph
  g <- d1 %>% 
    filter(!assessment == "total_score") %>% 
    ggplot(aes(x = assessment, y = score  ))+
    geom_point(shape = 21, fill = "salmon", color = "blue", size = 4)+
    geom_line(aes(group = student_id), linetype = "dashed")+ 
    geom_text(aes(label = score %>% round(1)), hjust = -1)+
    labs(
      title = paste0(
        "Course performance for student [", selected_student_id,"]","\n",
        "Number of exams missing = ", n_exams_missing, "; Total score = ", selected_total_score %>% round(1)
      )
    )+
    theme_bw()
  return(g)
}
# How to use
ds1 %>% graph_student_progress(1)
ds1 %>% graph_student_progress(25)
ds1 %>% graph_student_progress(19)
```
# Session Info
```{r}
sessionInfo()
```

