---
title: "Florida Suicide Trends"
author: "Andriy Koval"
date: "April 12, 2020"
output:
  html_document:
    theme: simplex
    toc: true
    toc_depth: 3
    keep_md: true
    toc_float: true
    code_folding: show
---

<!-- These two chunks should be added in the beginning of every .Rmd that you want to source an .R script -->
<!--  The 1st mandatory chunck  -->
<!--  Set the working directory to the repository's base directory -->
```{r, echo=F, message=F} 
#Don't combine this call with any other chunk 
# cat("Working directory: ", getwd()) # check where you are
  library(knitr)
# Rmd is in "./reports/reportA/reportA.Rmd", but now the root is "./"
  knitr::opts_knit$set(root.dir='../../') # "go up" from the folder where this .Rmd is save until you arrive at the root/working directory
```

<!--  The 2nd mandatory chunck  -->
<!-- Set the report-wide options, and point to the external code file. -->
```{r set_options, echo=F}
# set options shared by all chunks
opts_chunk$set(
  results='show', 
  attr.source = ".numberLines",
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
  fig.height = 4, 
  fig.width = 5.5, 
  out.width = "550px",
  fig.path = 'figure_rmd/',     
  dev = "png",
  dpi = 400
  # fig.path = 'figure_pdf/',     
  # dev = "pdf"#,
  # dev.args=list(pdf = list(colormodel = 'cmyk'))
)
echoChunks <- FALSE
options(width=120) #So the output is 50% wider than the default.
# connect to the file with the underlying R script  
read_chunk("./analysis/final-project/florida-suicides.R") # path from the root to the R script.
# read_chunk("../../../analysis/blogposts/florida-demographic-growth/fl-demo-growth.R") # path from the root to the R script. 
```

# Environment
<!-- Load packages to be used in the report  --> 
```{r load-packages, echo=TRUE, message=FALSE}
```

<!-- Load 'sourced' R files.  Suppress the output when loading packages. --> 
```{r load-sources, echo=TRUE, message=FALSE}
```

<!-- Load any Global functions and variables declared in the R file.  Suppress the output. -->
 ```{r declare-globals, echo=echoChunks, results='show', message=FALSE}
 ```
# Data

```{r load-data, echo=TRUE, results='show', message=FALSE}
ds0 %>% dplyr::glimpse(90)
```

# Tweaking

```{r tweak-data-1, echo=TRUE, results='show', message=FALSE}
```

# Population trends 

## 1. Total count 
> How does the total population of Florida changes between 2006 and 2020?

```{r population_1, echo=echoChunks, fig.height=5, fig.width=9, out.width = "900px",message=FALSE, results='show',cache=TRUE}
```

## 2. Within age groups
> What was the trajectory of population growth for each age group?

```{r population_2, echo=echoChunks, fig.height=5, fig.width=9, out.width = "900px",message=FALSE, results='show',cache=TRUE}
```

## 3. Sex and age groups
> For residends between 10 and 84 years of age, what was the trajectory of growth for each age group by sex? 

```{r population_3, echo=echoChunks, results='show', fig.height=5, fig.width=9,out.width = "900px", message=FALSE,cache=TRUE}
```

## 4. Ethnic group
> For residends between 10 and 84 years of age, what was the trajectory of growth for each ethnic group?

```{r population_4, echo=echoChunks, fig.height=7, fig.width=9, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```


# Suicide counts 

## 1. Total count 
> What is the trajectory of total suicides in FL between 2006 and 2017?

```{r suicide_1, echo=echoChunks, fig.height=5, fig.width=9, out.width = "900px",message=FALSE, results='show',cache=TRUE}
```

## 2. Men vs Women
> How does the trend of total suicides differ between men and women?

```{r suicide_2, echo=echoChunks, fig.height=5, fig.width=9, out.width = "900px",message=FALSE, results='show',cache=TRUE}
```

## 3. Sex and age groups
>For residends between 10 and 84 years of age, how does the trend of suicides counts among ethnic groups differ by sex?  

```{r suicide_3, echo=echoChunks, results='show', fig.height=5, fig.width=9,out.width = "900px", message=FALSE,cache=TRUE}
```

## 4. Ethnic group

> For residends between 10 and 84 years of age, how does the trend of total suicides between men and women differ across ethnic groups?  

```{r suicide_4, echo=echoChunks, fig.height=7, fig.width=9, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```


# Suicide rates

## 1. Function `compute_rates`

```{r compute_rate_function, echo=echoChunks, fig.height=7, fig.width=9, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

Compose the function that computes a rate of suicides per `100,000` for a given aggregation frame. To help you get started, here's the general shape of this function
```{r eval=F}

compute_suicide_rate <- function(d_input, group_by_variables){
  d_output <- d_input %>%

  return(d_output)
}
```
Add necessary line, so that the function behaves in the following way and generates the following output:   
```{r}
d <- ds0 %>% compute_suicide_rate(group_by_variables = c("year","sex")) # `group_by_variables` may be different
d %>% neat()
```

> Hint 1  

You will need to use `group_by` with `.dots` argument

> Hint 2

Make sure you `ungroup` before computing the rates in a separate `mutate` statement

## 2. Total rates
>  What is the trend of the total suicide rates in Florida between 2006 and 2017?  

```{r suicide_rate_1, echo=echoChunks, fig.height=7, fig.width=9, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

## 3. Youth by race and gender
> For residends between 10 and 24 years of age, how does the rate of suicide vary by race and gender?  

```{r suicide_rate_2, echo=echoChunks, fig.height=7, fig.width=9, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```


# Further insights

What other interesting patterns did you discern in this data? 


session information
===========================================================================

For the sake of documentation and reproducibility, the current report was rendered in the following environment.  Click the line below to expand.

<details>
  <summary>Environment <span class="glyphicon glyphicon-plus-sign"></span></summary>
```{r session-info, echo=FALSE}
if( requireNamespace("devtools", quietly = TRUE) ) {
  devtools::session_info()
} else {
  sessionInfo()
} 
```
