---
title: "Univariate Distributions"
author: "Andriy Koval"
date: "3/1/2020"
output: 
  html_document: 
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Agenda for today:  

1. Import and inspect `12 - Clinic Patient Demography` and `06 - Patient Readmission Days` data sets
2. Join two data sets using `dplyr::left_join()`
3. Create a univariate scatterplot with jittering
4. Describe distributions with a 5 number summary using `base::summary()`
5. Build a boxplot of patient age across race 
6. Build a violin plot of days to readmission across race

# Reference materials:
- Chapter 17 `Box-and-Whiskers Plot` in `Tableau for Healthcare` 3rd edition
- data source:  http://www.healthdataviz.com/resources/ 

## From `R Graphics Cookbook`:  
- 6.6 Making a Basic Box Plot (https://r-graphics.org/recipe-distribution-basic-boxplot )
- 6.8 Adding Means to a Box Plot (https://r-graphics.org/recipe-distribution-boxplot-mean)
- 6.9 Making a Violin Plot(https://r-graphics.org/recipe-distribution-violin)

## Other 
- Shapes for ggplot2 (http://sape.inf.usi.ch/quick-reference/ggplot2/shape) 
- Violin plots explained (https://towardsdatascience.com/violin-plots-explained-fb1d115e023d)

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(magrittr)
library(dplyr)
library(ggplot2)
library(explore)
```


# Load Data
```{r load-data}
path_file <- "../../data-public/raw/Tableau_10_Training_Files/Tableau 10 Training Practice Data.xlsx"
sheet_names <- readxl::excel_sheets(path_file)
dto <- list()
for(sheet_i in sheet_names){
   # i <- sheet_names[1]
  dto[[sheet_i]] <- readxl::read_xlsx(path_file, sheet = sheet_i)
}
ds_demography <- dto$`12 - Clinic Patients Demogr`
ds_readmissions <- dto$`06 - Patient Readmission Days`
```

## Inspect Patient Demography
```{r}
ds_demography %>% DT::datatable()
ds_demography %>% explore::describe_all()
```

## Inspect Readmissions
```{r}
ds_readmissions %>% DT::datatable()
ds_readmissions %>% explore::describe_all()

```

# Tweak data
## Demography
```{r tweak-data-demography}
names(ds_demography) <- c("patient_id", "gender", "race", "age")
ds_demography <- ds_demography %>% 
  dplyr::mutate(
    female = (gender == "F"),
    race  = forcats::fct_recode(race,
                                "Asian"     = "As"
                                ,"Black"    = "B"
                                ,"Hispanic" = "H"
                                ,"White"    = "W"
                                )
    
  ) %>% 
  dplyr::select(-gender) 
ds_demography %>% glimpse()

```

## Readmissions
```{r tweak-data-readmission}
names(ds_readmissions) <- c("patient_id", "cardio_patient", "days_to_readmission")
ds_readmissions <- ds_readmissions %>% dplyr::select(-cardio_patient)
ds_readmissions %>% glimpse()
```

## Join tables
```{r join}
ds <- dplyr::left_join(
  ds_demography
  ,ds_readmissions
  ,by = "patient_id"
)
ds %>% glimpse()

```

# Describe distribution
```{r}
ds %>% base::summary()
# ds %>% ggplot(aes(x = "Patients", y = age)) + geom_point()
# ds %>% ggplot(aes(x = "Patients", y = age)) + geom_jitter(width = .1)
ds %>% 
  ggplot(aes(x = race, y = age)) + 
  geom_boxplot()+
  geom_jitter(width = .2, alpha = .1, size = 3)+
  stat_summary(fun.y = mean, geom = "point", shape = 1, size = 4, color = "blue")+
  # stat_summary(fun.y = median, geom = "point", shape = 3, size = 5, color = "red" )+
  theme_bw()

```

## Boxplot: Age

```{r}
g0 <- ds %>% 
  ggplot(aes(x = race, y = age)) + 
  geom_boxplot(width = .5)+
  # geom_violin()+ # example of a violin
  geom_point(
    position = position_jitter(width = .2, height = 0, seed = 42 )
     ,shape = 21, color = "black", fill = "salmon",   alpha = .2, size = 4
    )+
  stat_summary(fun.y = mean, geom = "point", shape = 1, size = 4, color = "blue")+
  labs(
    title = "Age Distribution of Cardiovascular patients by Race",
    x = "Race", y = "Age in years"
  )+
  theme_bw()
g0
```

## Boxplot: Readmissions

```{r}
g1 <- ds %>% 
  ggplot(aes(x = race, y = days_to_readmission))+
  geom_violin( fill = "blue", alpha = .2, trim = F)+
  geom_boxplot(width = .1, fill = "grey90")+
  stat_summary(fun.y = mean, geom = "point", shape =21, fill = "white", size = 3)+
  labs(
    title = "Distribution of Days to Readmission among Cardiovascular Patients by Race"
    ,x = "Race", y = "Days to Readmission"
  )+
  theme_bw()
g1
  


```



